# Nelson Muteti : Independent Project Week 9

## 1.0 Defining the Question
**A Kenyan entrepreneur has created an online cryptography course and wants to advertise it on her blog.**
**She currently targets audiences originating from various countries. In the past, she ran ads to advertise**
**a related course on the same blog and collected data in the process. She would now like to employ my** **services as a Data Science Consultant to help her identify which individuals are most likely to click**
**on her ads**


## 1.1 Defining Metrics for Success

**A succesful project will clean the data provided, perform Univariate and Bivariate**
**analysis of the variables and finally draw recommendations for the business.**

## 1.3 Understanding The context.

**The context here is an advertising one. We need to first understand how to measure the success rate of ads**
**posted on a site**.

## 1.4 Reading the data

```{r}
# first imports library data.table

library(data.table)
library(rmarkdown)
```
```{r}
#fetch the data using fread()

df = fread('http://bit.ly/IPAdvertisingData')
```
```R
#output and get a glimpse of data

head(df)
```
```{r,message=FALSE}
#get the dimensions of the dataset

dim(df)

#data has 1000 rows and 10 columns

```

### 1.5 External Data Validation

**I ensured that the data provided was logical and data types given matched the values in columns**

```{r}
#get summary statistics of the data

summary(df)

```
```{r}
#get data types of each column

str(df)

```

**The target column is the "Clicked Ad " column.**

**1 - Clicked the ad.**
**0 - did not click the ad.**

### Recode the factor variables

```{r}
df$`Clicked on Ad` <-factor(df$`Clicked on Ad`, levels = c(0, 1), labels = c('No click', 'Clicked'))
df$Male <-factor(df$Male, levels = c(0, 1), labels = c('Female', 'Male'))

#convert the Male and Clicked Ad columns to factors

df$Male <- as.factor(df$Male)

#identify levels of column Male

levels(df$Male)
```


**Do the same for our target column**
```{r}

df$`Clicked on Ad` <- as.factor(df$`Clicked on Ad`)

# identify levels of target column

levels(df$`Clicked on Ad`)

```
```{r}
#now check the summary of columns

summary(df$`Clicked on Ad`)

# 500 people clicked on the ad while 500 did not click on the ad

```
## 1.6 Data Cleaning

**First, we will check for NA values in data.**

```R
# Now to check for NA values

is.na(df) # outputs boolean / logic values. True when there is NA values 

#           and False when there isn't

```
```{r}

sum(is.na(df)) # checking for total count of NA values

# the count is 0;meaning no missing values in data

# now for a closer inspection of columns and any missing values 

# that may have been skipped

```
```R
colSums(is.na(df))

# 0 missing values in each individual column

```

**Next we check for duplicated rows of data.**

```{r}

#we call the unique rows mydf

mydf <- unique(df)

#check if rows were dropped

dim(mydf)

# no rows were dropped since all unique values correspond to initial dataset dimensions

```

**Next we identify outliers in Numerical columns.**

```{r}

# first identify the numerical columns

str(mydf)

# Numerical Columns : Daily Time Spent on Site,Age, Area Income, Daily Internet Usage

```

**Use a boxplot to identify outliers.**
```{r}

boxplot(mydf$`Daily Time Spent on Site`,col = 'blue',main="Checking for outliers in Time spent on Site",xlab="Daily Time Spent on Site")

# data points in this column seem to be well distributed
```

```{r}
#Now lets get all the outlier data points from the column

boxplot.stats(mydf$`Daily Time Spent on Site`)$out

# there are 0 outliers here
```

**We do the same for the other two numerical columns.**

```{r}

boxplot(mydf$Age,col = 'red',main="Checking For Outliers in Age",xlab="Age")

boxplot.stats(mydf$Age)$out

```
```{r}

boxplot(mydf$`Area Income`,col = 'green',main="Checking for outliers in Area Income",xlab="Area Income")

boxplot.stats(mydf$`Area Income`)$out

length(boxplot.stats(mydf$`Area Income`)$out)

```

**8 outliers in the column.These few outliers will not have any effect on**

**the data. They are important to the data.**
```{r}

boxplot(mydf$`Daily Internet Usage`,col = 'yellow',main="Checking for outliers in Internet Usage",xlab="Daily Internet Usage")

boxplot.stats(mydf$`Daily Internet Usage`)$out

```
## 1.7 Univariate EDA

**Now we proceed to Univariate EDA.**
```{r}

# first get measures of central tendency

mean(mydf$`Daily Time Spent on Site`)

median(mydf$`Daily Time Spent on Site`)

```

```{r}
# function to get mode of data

mymod <- function(v){uniqv <- unique(v)
uniqv[which.max(tabulate(match(v,uniqv)))]

}

```

```{r}
mean(mydf$Age)

# mean age of customers visiting site is 36
```

```{r}
quantile(mydf$Age)

# age distribution of customers on site by quantiles
```

```{r}
mean(mydf$`Area Income`)

quantile(mydf$`Area Income`)

# majority of customers visiting the site are from the  middle class 

```
```{r}

mymod(mydf$`Ad Topic Line`) # identify the most advertised topic

#Cloned 5thgeneration orchestration was the most advertised Ad to customers visiting the site

```
```{r}
mymod(mydf$City) # which is the most popular city of site visitors

# Lisamouth is the most popular city 
```
```{r}

sapply(mydf, function(x) length(unique(x))) # get unique values of each column

#so there are 237 unique countries in the dataFrame

```

**Barplot for Gender in data.**

```{r}

df_gender <- mydf$Male

df_gender <- table(df_gender)

barplot(df_gender,col = 'blue',xlab = " Gender",ylab = "Count",main = "Gender Count in Data")

```

**Barplot for Ad clicks in data**

```{r}

df_click <- mydf$`Clicked on Ad`

df_click <- table(df_click)

barplot(df_click,col = 'red',xlab = " Clicked Ad ?",ylab = "Count",main = "Ad Click Count in Data")
```



**Age histogram**


```{r}

hist(mydf$Age,col = 'green',xlab = 'Ages',main = "Age Distribution in data",freq = FALSE)

# include a Kernel density estimator line on Histogram

dens <- density(mydf$Age)
# include the KDE line
lines(dens)

```
```{r}

hist(mydf$`Daily Time Spent on Site`,col = 'red',xlab = 'Time Spend On Site',main = "Distribution Of Time Spend on site",freq = FALSE)

# include a Kernel density estimator line on Histogram

dens <- density(mydf$`Daily Time Spent on Site`)
# include the KDE line
lines(dens)
```


**Distribution of ages in data seems to be normal (bell - curve).**



**Now lets get the ranges,standard Deviations, Variances  and skewness of various** **Numerical columns**

```{r}
#range of Age

range(mydf$Age)

```
```{r}
#variance of age

var(mydf$Age)

```
```{r}
# range of incomes

range(mydf$`Area Income`)

```
```{r}
# variance of incomes in data

var(mydf$`Area Income`)

```
```{r}

library(moments)
```


Now check for skewness and kurtosis of time spend on site and Internet usage columns


```{r}

skewness(mydf$`Daily Time Spent on Site`)

```

**Time spend on site has a negative skewness. This means that majority ****of data  points****are greater than the mean.On average, most people spend a lot of time on the** **site.**


```{r}

skewness(mydf$`Daily Internet Usage`)

```


**1.Daily Internet usage also has a negative skewness. This means that majority ****of data points****are greater than the mean.Most people visiting the site use more than the average daily internet**


## 1.8.Bi-Variate Analysis

**First, a side by side boxplot.**
```{r}

boxplot(Age ~ Male,data = mydf,main = "Ages by Gender")

#the distribution of ages does not vary with gender
```
```{r}
tapply(mydf$Age,mydf$Male,mean)

# get the mean of ages grouped by gender..as we can see, mean does not vary with gender
```


**Get a side by side boxplot for ad clicks.**

```{r}
boxplot(`Daily Time Spent on Site`~`Clicked on Ad`,data = mydf,main = "Time On Site Against Clicking Ads")

# it seems that people who spend less time on the site are more likely to click on

# the ads
```
```{r}

# Identify difference in times of people who either clicked or did not click ADs

tapply(mydf$`Daily Time Spent on Site`,mydf$`Clicked on Ad`,mean)

# the average time of people who clicked on ADs is considerably less than

# those who did not click on ADs
```

```{r}

boxplot(`Daily Internet Usage`~`Clicked on Ad`,data = mydf,main = "Internet Usage vs Clicking on Ads")

# people with less daily internet usage were more likely to click an Ad
```
```{r}

tapply(mydf$`Daily Internet Usage`,mydf$`Clicked on Ad`,mean)

```


**Distribution of age vs Income**

```{r}

boxplot(Age~`Area Income`,data = mydf,main = "Age Against Area Income")
```



**Find the correlation between age and are income.**

```{r}

cor(mydf$Age,mydf$`Area Income`)

# weak negative correlation between age and income
```



**Distribution of time spend on site and daily internet usage.**
```{r}

plot(`Daily Time Spent on Site`~`Daily Internet Usage`,data = mydf,main = "Time Spend On Site Against Daily Internet Usage Of Customers")
```



Now find corrrelation between daily time spend on sight and internet usage
```R

cor(mydf$`Daily Time Spent on Site`,mydf$`Daily Internet Usage`)

# there is a relatively moderate correlation between time spend on site and internet usage
```


**Now lets plot all scatter pairs for numeric variables.**

```{r}

mydf$Age <- as.numeric(mydf$Age)

pairs(mydf[,-c(5,6,7,8,9,10)])

# now we have all scatter plot pairs from the data

```
```{r}

library(ggplot2)
```
```{r}
ggplot(mydf,
       aes(x = `Clicked on Ad`,
           fill = Male))+
  geom_bar(position = "dodge")+
  labs(title = " Analysis of Ad Clicking by Gender")
```



## 1.9 .Multivariate analysis


**Get the corrrelation matrix from the data.**
```{r}

cor(mydf[,-c(5,6,7,8,9,10)])

# we can see that the area income and daily internet usage have a weak but positive

# linear relationship

# the daily time spend on site is also moderately correlated to the daily internet use
```


**Now lets get the multivariate aggregates of numerical variables.**

```{r}

aggregate(mydf$`Daily Time Spent on Site`~mydf$`Clicked on Ad`+mydf$Male,FUN = mean)
```


**Now lets get the number of people that clicked on ads given their gender only using a crosstab.**

```{r}

table(mydf$`Clicked on Ad`,mydf$Male)

#relatively more females clicked on ADs. But this could be due to the reason that there were more women sampled
```
```{r}

library(pch)
```


**Now lets investigate the relationship between clicking of ads and internet usage + time spend on site.**

```{r}

ggplot(mydf,
       aes(x = `Daily Time Spent on Site`,
           y = `Daily Internet Usage`,
           color=`Clicked on Ad`))+
  geom_point()+labs(title = "Time spend on site vs Internet usage on Clicking Ads")

```


## 2.0. Recommendations.


**1.We can see that individuals who spend less time on the site and had less data per day were more likely to** **click on the ADs and had less data per day were more likely to click on the ADs.The Ad company should most** **probably target these individuals.** 


**2.Since most of the customers visiting the site spend more time on the site than the** **average time spent, the Advertising team should consider pitching Ads that are** **relevant to the site or topic they are interested in.**

## 2.1. Challenging The Solution.

**The data acquired should probably have been more to do with the topics of the site**
**visited rather than the clients visiting the site.**

**This would have given has more clarity on the type of interests the customers have.**

